---
title: "Data import and cleaning"
---

This notebook is the starting point for the power plants data project. I aim to focus on how the location of power plants in Texas intersect with the community and population that pre-exist in the area. Demographic data will be pulled from the [Texas Demographics Center](https://demographics.texas.gov/Estimates/#FAQ) (specifically age, sex and race/ethnicity for state and counties) and [US Department of Agriculture](https://data.ers.usda.gov/reports.aspx?ID=4038).

## Setup

Setting up the packages that will be used in this notebook.

```{r}
#| message: FALSE

library(tidyverse)
library(janitor)
library(tidycensus)
library(sf)
```

Activating my Census API key in case I wanted to use Census data for any of the things I am looking for. 

```{r}
census_api_key("460e8dedb542f01b4ea4adede880098417e04166")
```


## Importing files

### Race

Starting off with race data, taken from TDC.

```{r}
race_pop_raw <-
  read_csv("demographic-data/race-pop.csv")

race_pop_raw
```

I am only interested in the total of each demographic group in the counties. I also want the most recent data from 2024, but do note that they are estimates so they are not as accurate as the Census data would be.

### Income

Now importing median household income from USDA. Noting that I edited this file in Excel myself. I got rid of style additions, spaces and fixed the county column to fit R's needs.


```{r}
income_raw <-
  read_csv("demographic-data/UnemploymentReport.csv")

income_raw
```

I am only interested in the median income and the counties, so I will be cleaning for that. These two rows are also characters instead of numeric, so I will need to fix that before doing anything with the data set. The years in the data above are unemployment rates since this data is taken from the unemployment data for Texas.

## Cleaning up through filter and select

### Race

The race population data is quite large, so I am going to filter the data for what I want (total of ethnic/racial groups, counties) and see how to make it more digestible. Note this data is from July 2024 and these are estimates for the population.

```{r}
race_pop_cleaned <- race_pop_raw |> 
    filter(Age == "All Ages",
           Year == "202407") |> 
  select(Areaname,
         Total,
         NH_White_Total,
         NH_Black_Total,
         NH_Asian_Total,
         NH_Other_Total,
         Hispanic_Total) |> 
  rename("Area" = "Areaname",
         "NH_White" = "NH_White_Total",
         "NH_Black" = "NH_Black_Total",
         "NH_Asian" = "NH_Asian_Total",
         "NH_Other" = "NH_Other_Total",
         "Hispanic"= "Hispanic_Total") |>  filter(!Area == "TEXAS")

race_pop_cleaned
```



### Income

Now that our race population is clean, we are gonna tackle our median income data. Note that the median household income is from 2022.


```{r}
income_cleaned <- income_raw |> 
  select(Area,
         median_house_income_22,
         perc_of_state_income) |> 
  rename("Median_HH_Income" = "median_house_income_22",
         "State_HH_Inc_Percent" = "perc_of_state_income") |> 
  filter(!Area == "TEXAS")

income_cleaned
```

## Getting geometry to use for mapping

Getting the data needed to successfully map this. I found this through [Texas DoT](https://gis-txdot.opendata.arcgis.com/datasets/9b2eb7d232584572ad53bad41c76b04d_0/explore?location=31.119954%2C-100.049436%2C5).

```{r}
counties_map <-
  read_sf("mapping-data/Texas_County_Boundaries_-2028607862104916578.geojson") |> 
  arrange(CNTY_NM) |> 
  select(CNTY_NM,
         geometry) |> 
  rename("Area" = "CNTY_NM") |> 
  mutate(Area = str_c(Area, "County", sep = " "),
         Area = toupper(Area))

counties_map
```


## Joining data

Joining my demographic data with the spatial data I found.

### Race

Starting with race.

```{r}
race_counties_joined <- race_pop_cleaned |> 
  inner_join(counties_map, by = "Area")

race_counties_joined
```

Turning the race joined data into an SF object.

```{r}
race_counties_joined |> 
  write_sf("interactive-data/race_counties.geojson")
```

### Misc

Turning the counties data into an SF object for a mapping objective I was trying to achieve.

```{r}
counties_map |> 
  write_sf("interactive-data/counties_map.geojson")
```

### Income

Joining income data with the location data.

```{r}
income_counties_joined <- income_cleaned |> 
  inner_join(counties_map, by = "Area")

income_counties_joined
```

Turning into an SF object.

```{r}
income_counties_joined |> 
  write_sf("interactive-data/income_counties.geojson")
```

## Plant locations

Found a database of plant locations. I will import and clean it up here.

### Importing files

These were compiled from [Oil and Gas Watch](https://oilandgaswatch.org/summary?facilitiesAndOrPipelines=Facilities+Only&state=TX). I downloaded the database, but it includes all states and several miscellaneous pieces of information.

```{r}
plant_raw_data <- read_csv("plant-data/plant-projects-all.csv")

plant_raw_data
```

### Cleaning plant data

Now that it is imported, we can clean it up for the data I want.

```{r}
plant_clean <- plant_raw_data |> 
  filter(State == "TX",
         !Latitude == "NA",
         !Longitude == "NA") |> 
  select(Area,
         Company,
         Project,
         Status,
         Classification,
         Industry,
         Type,
         Latitude,
         Longitude) |> 
  mutate(Area = toupper(Area)) |> 
  arrange(Area)

plant_clean
```

Note: I noticed that some entries have multiple counties, so I went and cleaned them in OpenRefine.

### Export

Exporting to turn into a sf object.

```{r}
plant_clean |> 
  write_csv("interactive-data/tx-plant-data.csv")
```

Also turning it into an rds for an objective I was trying to achieve in mapping.

```{r}
plant_clean |> 
  write_rds("interactive-data/tx-plants-data.rds")
```


### SF object and exporting

```{r}
tx_plant_data <- read_csv("interactive-data/tx-plant-data.csv")
```

```{r}
tx_plant_data_sf <- tx_plant_data |> 
  st_as_sf(
    coords = c("Longitude", "Latitude"),
    crs = 4326
  )
```

Now, I am exporting it.

```{r}
tx_plant_data_sf |> write_sf("interactive-data/tx-plants-data.geojson")
```

## Exporting as rds

Exporting the race and income data as rds to use in my analysis.

```{r}
race_pop_cleaned |> 
  write_rds("demographic-data/tx-race-data.rds")
```

```{r}
income_cleaned |> 
  write_rds("demographic-data/tx-income-data.rds")
```

## Importing pollution data

Importing and cleaning data I will be using for analysis.

Below are air quality readings sourced from [EPA](https://www.epa.gov/outdoor-air-quality-data/air-quality-index-report
) for the years 2025, 2015, 2005.

```{r}
air_pol1 <- read.csv("pol-data/aqireport2025.csv")

air_pol1
```

```{r}
air_pol2 <- read.csv("pol-data/aqireport2015.csv")
```

```{r}
air_pol3 <- read.csv("pol-data/aqireport2005.csv")
```

## Cleaning pollution data

There are two things I want to do before filtering and renaming: clean out the ", TX" and make them all caps. I will do that here, then select and rename.

```{r}
air_pol_25 <- air_pol1 |> 
  mutate(County = str_remove(County, ", TX"),
         County = toupper(County)) |> 
  select(County,
         AQI.Maximum,
         AQI.Median) |> 
  rename("Max_AQI" = "AQI.Maximum",
         "Median_AQI" = "AQI.Median")

air_pol_25
```

```{r}
air_pol_15 <- air_pol2 |> 
  mutate(County = str_remove(County, ", TX"),
         County = toupper(County)) |> 
  select(County,
         AQI.Maximum,
         AQI.Median) |> 
  rename("Max_AQI" = "AQI.Maximum",
         "Median_AQI" = "AQI.Median")

air_pol_15
```

```{r}
air_pol_05 <- air_pol3 |> 
  mutate(County = toupper(County)) |> 
  select(County,
         AQI.Maximum,
         AQI.Median) |> 
  rename("Max_AQI" = "AQI.Maximum",
         "Median_AQI" = "AQI.Median")

air_pol_05
```

## Exporting pollution data

Exporting it all to use in analysis notebook.

```{r}
air_pol_25 |> write_rds("pol-data/air_pol_25")
```

```{r}
air_pol_15 |> write_rds("pol-data/air_pol_15")
```

```{r}
air_pol_05 |> write_rds("pol-data/air_pol_05")
```