---
title: "Asian population maps"
format: html
---

## Goals of this notebook

I want to map the demographic data about the Texas Asian population I found and cleaned in the first notebook. 

## Setup

I am setting up the packages I need.

```{r}
library(leaflet)
library(tidyverse)
library(tidycensus)
library(janitor)
library(tigris)
library(scales)
library(sf)
library(tmap)
library(shiny)
```

## Getting the map started

Importing the data that I cleaned in the cleaning notebook.

```{r}
counties_map <- read_sf("interactive-data/race_counties.geojson")

counties_map
```

## Demographic maps

### Percentages

Starting out with making a percentage of the Asian population in Texas.

```{r}
asian_map <- counties_map |> 
  mutate(nh_asian_pct = NH_Asian / Total) |>
  select(Area, nh_asian_pct)

asian_map
```

### Static 

Making the static map.

```{r}
asian_map |> 
  shift_geometry() |> 
  ggplot() +
  geom_sf(
    aes(
      fill = nh_asian_pct,
      color = nh_asian_pct
    )
  ) +
  theme_void() +
  theme(
    legend.position = "top",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.5, "cm"),
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16,
      margin = margin(
        b =0.25,
        unit = "cm"
      )
    )
  ) +
  scale_fill_viridis_c(limits = c(0,1),
                       labels = percent_format()) +
scale_color_viridis_c(limits = c(0,1),
                     labels = percent_format()) +
  labs(
    fill = NULL,
    color = NULL,
    title = "Percentage of Asian Population Estimates for Texas Counties"
  )
```


### Interactive

Making an interactive version.

Starting with making the labels.

```{r}
asian_map_labels <- asian_map |> 
  mutate(text_label = str_glue(
    "<b>{percent(nh_asian_pct, accuracy = 0.1)}</b> of the population of <b>{Area}</b> identifies as Asian"
  ))
```


Making a color palette.

```{r}
my_palette_2 <- colorNumeric(
  palette = "viridis",
  domain = asian_map_labels$nh_asian_pct
)
```


Creating the map itself.

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = asian_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_2(nh_asian_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_2,
    values = asian_map_labels$nh_asian_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```

## Population with plant data

### Plant data

Getting the plant map here first.

```{r}
plants_map <- read_sf("interactive-data/tx-plants-data.geojson")

plants_map
```

Then making the labels.

```{r}
plants_map_labels <- plants_map |> 
  mutate(text_label = str_glue(
    "{Project} developed by {Company}"
  ))
```

### Map

Now making the map.

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = asian_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_2(nh_asian_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_2,
    values = asian_map_labels$nh_asian_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  ) |> 
  addCircleMarkers(
    data = plants_map_labels,
    fillColor = "red",
    stroke = FALSE,
    radius = 3,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```