---
title: "Income in Texas map"
format: html
---

## Goals of this notebook

I want to map the demographic data about the Texans income I found and cleaned in the first notebook. 

## Setup

I am setting up the packages I need.

```{r}
library(leaflet)
library(tidyverse)
library(tidycensus)
library(janitor)
library(tigris)
library(scales)
library(sf)
library(tmap)
library(shiny)
```

## Getting the map started

Importing the data that I cleaned in the cleaning notebook.

```{r}
counties_map <- read_sf("interactive-data/race_counties.geojson")

counties_map
```
Importing the median household income per county.

```{r}
tx_income <- read_sf("interactive-data/income_counties.geojson")
```

I know that some of the data is character instead of a number, so I am fixing that here.

```{r}
income <- tx_income |> 
  mutate(across(c(Median_HH_Income, State_HH_Inc_Percent), parse_number),
         HH_Inc_Perc = State_HH_Inc_Percent / 100)

income
```

## Interactive map

Now creating the labels for the map.

```{r}
income_labels <- income |> 
  mutate(text_label = str_glue(
    "The median income in {Area} is {Median_HH_Income}"
  ))
```

Creating the color palette now.

```{r}
my_palette_5 <- colorNumeric(
  palette = "viridis",
  domain = income_labels$HH_Inc_Perc
)
```

Making an interactive map.

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = income_labels,
    weight = 0.75,
    fillColor = ~my_palette_5(HH_Inc_Perc),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_5,
    values = income_labels$HH_Inc_Perc,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```

Note: The percentages in this map refer to how much of Texas' median income a particular county is close to. For example, keeping in mind the median income in Texas is $72,279, Travis county is around 120% of the Texas median income, at $95,151. 

## Income with plants

### Plant data

Getting the plant map here first.

```{r}
plants_map <- read_sf("interactive-data/tx-plants-data.geojson")

plants_map
```

Then making the labels.

```{r}
plants_map_labels <- plants_map |> 
  mutate(text_label = str_glue(
    "{Project} developed by {Company}"
  ))
```

### Map

Now making the map.

```{r}
leaflet() |> 
    addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = income_labels,
    weight = 0.75,
    fillColor = ~my_palette_5(HH_Inc_Perc),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_5,
    values = income_labels$HH_Inc_Perc,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  ) |> 
  addCircleMarkers(
    data = plants_map_labels,
    fillColor = "red",
    stroke = FALSE,
    radius = 3,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```


