---
title: "Texas power plants map"
format: html
---

## Goals of this notebook

I want to map the plant data I found and cleaned in the first notebook. 

## Setup

I am setting up the packages I need.

```{r}
library(leaflet)
library(tidyverse)
library(tidycensus)
library(janitor)
library(tigris)
library(scales)
library(sf)
library(tmap)
library(shiny)
```

## Getting the map started

Importing the data that I cleaned in the cleaning notebook.

```{r}
counties_map <- read_sf("interactive-data/race_counties.geojson")

counties_map
```


## Plant data

Mapping the plant data as locations.

Starting out with a static map.

```{r}
plants_map <- read_sf("interactive-data/tx-plants-data.geojson")

plants_map
```

Using a package to import the geometry of the state of Texas to create the base of the static map.

```{r}
texas <- map_data('county', region = 'texas')

texas
```

Importing the rds version of the Texas plants.

```{r}
tx_plants <- read_rds("interactive-data/tx-plants-data.rds")
```

## Static map

Creating the static map.

Starting out with the base layer.

```{r}
texas_map <- ggplot(data = texas) +
  geom_polygon(aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  coord_fixed(1.3)

texas_map
```

Putting the plant locations on the map.

```{r}
texas_map +
  geom_point(data = tx_plants, aes(x = Longitude, y = Latitude), color = "blue", size = 0.75) +
  theme_void()
```

## Interactive map

Done with static, making an interactive version.

Creating labels for the map.

```{r}
plants_map_labels <- plants_map |> 
  mutate(text_label = str_glue(
    "{Project} developed by {Company}"
  ))
```

Creating the interactive map.

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addCircleMarkers(
    data = plants_map_labels,
    fillColor = "blue",
    stroke = FALSE,
    radius = 3,
    fillOpacity = 1,
    popup = ~text_label
  ) 

```



