---
title: "Data mapping"
---

## Goals of this notebook

I want to map the demographic data we found and cleaned in the first notebook. This will adjust as time goes on, but this is the main focus for now.

I am following a tutorial from [here](https://rfortherestofus.com/courses/mapping/lessons/welcome-to-mapping-with-r).

## Setup

I am setting up the packages I need.

```{r}
library(leaflet)
library(tidyverse)
library(tidycensus)
library(janitor)
library(tigris)
library(scales)
library(sf)
library(tmap)
library(shiny)
```

## Getting the map started

Importing the data.

```{r}
counties_map <- read_sf("interactive-data/race_counties.geojson")

counties_map
```

```{r}
counties_map_wgs84 <-
  counties_map |> st_transform(crs = 4326)
```


```{r}
# leaflet() |> 
#   addTiles("CartoDB.Positron", group = "Map") |> 
#   addPolygons(data = counties_map_wgs84,
#               weight = 0.2,
#               color = "purple",
#               group = "Race") |> 
#   addLayersControl(
#     baseGroups = "Map",
#     overlayGroups = "Race"
#   )
```


```{r}
```


```{r}
```

## Non-white category

Going to create a category that adds up Black, Asian and Hispanic populations and make that into a choropelth map.

```{r}
nw_counties_map <- counties_map_wgs84 |> 
  mutate(BIPOC_Pop = NH_Black + NH_Asian + NH_Other + Hispanic)

nw_counties_map
```

Now making the map.

```{r}
# my_palette <-
#   colorNumeric(
#     palette = "viridis",
#     domain = nw_counties_map$BIPOC_Pop
#   )
```

```{r}
# leaflet() |> 
#   addProviderTiles("CartoDB.Positron") |> 
#   addPolygons(
#     data = nw_counties_map,
#     fillColor = ~my_palette(BIPOC_Pop),
#     fillOpacity = 0.8,
#     color = "white",
#     weight = 0.75,
#     label = ~BIPOC_Pop
#   )
```

Not sure how much I like that, so I will come back to it after trying other options. First, I want to try adding labels, though.

```{r}
nw_counties_map_labels <- nw_counties_map |>
  mutate(
    text_label = str_glue(
      "{BIPOC_Pop} of {Total} are considered as BIPOC in {Area}"
    )
  )
```
Making it into a map.

```{r}
# leaflet() |> 
#   addProviderTiles("CartoDB.Positron") |> 
#   addPolygons(
#     data = nw_counties_map_labels,
#     fillColor = ~my_palette(BIPOC_Pop),
#     fillOpacity = 0.8,
#     color = "white",
#     weight = 0.75,
#     popup = ~text_label
#   )
```

I don't like that either. Solution is to make it into a percentage, but I would have to double check what classifies as other first before going ahead. For now, I'm moving on.

## Dot/bubble map

Creating a bubble map.

```{r}
# leaflet() |> 
#   addProviderTiles("CartoDB.Positron") |> 
#   addCircles(
#     data = nw_counties_map,
#     label = ~BIPOC_Pop,
#     stroke = FALSE,
#     fillOpacity = 0.5,
#     radius = ~ Total / 100
#   )
```

Can't find a reason for the error, will come back to this later.

## Interactive map

Using [this](https://walker-data.com/census-r/mapping-census-data-with-r.html#choropleth-maps-with-tmap) as a guide.

```{r}
# groups <- nw_counties_map |> 
#   c("Hispanic", "NH_White", "NH_Black", "NH_Asian", "NH_Other")
# ```
# 
# 
# ```{r}
# dropdown <- fluidPage(
#   sidebarLayout(
#     sidebarPanel(
#       selectInput(
#         inputId = "group",
#         label = "Select a group",
#         chpices = groups
#       )
#     ),
#     mainPanel(
#       leafletOutput("map", height = "600")
#     )
#   )
# )
```

## Four different maps

Making four different interactive maps for the different populations. Important note that the races are mutually exclusive groups and non-Hispanic other consists of individuals such as Alaska and American Natives, Pacific Islanders and all people listing two or more races, according to the Texas Demographic Center.

### Black population

Starting with Black population of Texas.

```{r}
black_map <- counties_map |> 
  mutate(nh_black_pct = NH_Black / Total) |>
  select(Area, nh_black_pct)

black_map
```

Making the map.

```{r}
black_map |> 
  shift_geometry() |> 
  ggplot() +
  geom_sf(
    aes(
      fill = nh_black_pct,
      color = nh_black_pct
    )
  ) +
  theme_void() +
  theme(
    legend.position = "top",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.5, "cm"),
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16,
      margin = margin(
        b =0.25,
        unit = "cm"
      )
    )
  ) +
  scale_fill_viridis_c(limits = c(0,1),
                       labels = percent_format()) +
scale_color_viridis_c(limits = c(0,1),
                     labels = percent_format()) +
  labs(
    fill = NULL,
    color = NULL,
    title = "Percentage of Black Population Estimates for Texas Counties"
  )
```

Making an interactive version.

```{r}
black_map_labels <- black_map |> 
  mutate(text_label = str_glue(
    "<b>{percent(nh_black_pct, accuracy = 0.1)}</b> of the population of <b>{Area}</b> identifies as Black"
  ))
```


```{r}
my_palette_1 <- colorNumeric(
  palette = "viridis",
  domain = black_map_labels$nh_black_pct
)
```

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = black_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_1(nh_black_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_1,
    values = black_map_labels$nh_black_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```


### Asian population

Now, doing the same for Asian population.

```{r}
asian_map <- counties_map |> 
  mutate(nh_asian_pct = NH_Asian / Total) |>
  select(Area, nh_asian_pct)

asian_map
```

Making the map.

```{r}
asian_map |> 
  shift_geometry() |> 
  ggplot() +
  geom_sf(
    aes(
      fill = nh_asian_pct,
      color = nh_asian_pct
    )
  ) +
  theme_void() +
  theme(
    legend.position = "top",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.5, "cm"),
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16,
      margin = margin(
        b =0.25,
        unit = "cm"
      )
    )
  ) +
  scale_fill_viridis_c(limits = c(0,1),
                       labels = percent_format()) +
scale_color_viridis_c(limits = c(0,1),
                     labels = percent_format()) +
  labs(
    fill = NULL,
    color = NULL,
    title = "Percentage of Asian Population Estimates for Texas Counties"
  )
```

Making an interactive version.

```{r}
asian_map_labels <- asian_map |> 
  mutate(text_label = str_glue(
    "<b>{percent(nh_asian_pct, accuracy = 0.1)}</b> of the population of <b>{Area}</b> identifies as Asian"
  ))
```


```{r}
my_palette_2 <- colorNumeric(
  palette = "viridis",
  domain = asian_map_labels$nh_asian_pct
)
```

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = asian_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_2(nh_asian_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_2,
    values = asian_map_labels$nh_asian_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```

### Hispanic

Applying same Process for Hispanic population.

```{r}
hispanic_map <- counties_map |> 
  mutate(hispanic_pct = Hispanic / Total) |>
  select(Area, hispanic_pct)

hispanic_map
```

Making the map.

```{r}
hispanic_map |> 
  shift_geometry() |> 
  ggplot() +
  geom_sf(
    aes(
      fill = hispanic_pct,
      color = hispanic_pct
    )
  ) +
  theme_void() +
  theme(
    legend.position = "top",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.5, "cm"),
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16,
      margin = margin(
        b =0.25,
        unit = "cm"
      )
    )
  ) +
  scale_fill_viridis_c(limits = c(0,1),
                       labels = percent_format()) +
scale_color_viridis_c(limits = c(0,1),
                     labels = percent_format()) +
  labs(
    fill = NULL,
    color = NULL,
    title = "Percentage of Hispanic Population Estimates for Texas Counties"
  )
```

Making an interactive version.

```{r}
hispanic_map_labels <- hispanic_map |> 
  mutate(text_label = str_glue(
    "<b>{percent(hispanic_pct, accuracy = 0.1)}</b> of the population of <b>{Area}</b> identifies as Hispanic"
  ))
```


```{r}
my_palette_3 <- colorNumeric(
  palette = "viridis",
  domain = hispanic_map_labels$hispanic_pct
)
```

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = hispanic_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_3(hispanic_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_3,
    values = hispanic_map_labels$hispanic_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```

### Other

Finishing it off with other population, which, as mentioned above, includes Native Americans and people with multiple races.

```{r}
other_map <- counties_map |> 
  mutate(other_pct = NH_Other / Total) |>
  select(Area, other_pct)

other_map
```

Making the map.

```{r}
other_map |> 
  shift_geometry() |> 
  ggplot() +
  geom_sf(
    aes(
      fill = other_pct,
      color = other_pct
    )
  ) +
  theme_void() +
  theme(
    legend.position = "top",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.5, "cm"),
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16,
      margin = margin(
        b =0.25,
        unit = "cm"
      )
    )
  ) +
  scale_fill_viridis_c(limits = c(0,1),
                       labels = percent_format()) +
scale_color_viridis_c(limits = c(0,1),
                     labels = percent_format()) +
  labs(
    fill = NULL,
    color = NULL,
    title = "Percentage of Individuals Identifying as Other Estimates for Texas Counties"
  )
```

Making an interactive version.

```{r}
other_map_labels <- other_map |> 
  mutate(text_label = str_glue(
    "<b>{percent(other_pct, accuracy = 0.1)}</b> of the population of <b>{Area}</b> identifies as Other"
  ))
```


```{r}
my_palette_4 <- colorNumeric(
  palette = "viridis",
  domain = other_map_labels$other_pct
)
```

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = other_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_4(other_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_4,
    values = other_map_labels$other_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```

## Plant data

Mapping the plant data as locations.

Starting out with a static map.

```{r}
plants_map <- read_sf("interactive-data/tx-plants-data.geojson")

plants_map
```

```{r}
texas <- map_data('county', region = 'texas')

texas
```

```{r}
tx_plants <- read_rds("interactive-data/tx-plants-data.rds")
```

```{r}
texas_map <- ggplot(data = texas) +
  geom_polygon(aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  coord_fixed(1.3)

texas_map
```

```{r}
texas_map +
  geom_point(data = tx_plants, aes(x = Longitude, y = Latitude), color = "blue", size = 0.75) +
  theme_void()
```

Done with static, making an interactive version.

```{r}
plants_map_labels <- plants_map |> 
  mutate(text_label = str_glue(
    "{Project} developed by {Company}"
  ))
```


```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addCircles(
    data = plants_map_labels,
    fillColor = "blue",
    stroke = FALSE,
    radius = 500,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```

## Adding both sets of data on a map

### Black population

```{r}
leaflet() |> 
   addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = black_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_1(nh_black_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_1,
    values = black_map_labels$nh_black_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  ) |> 
  addCircles(
    data = plants_map_labels,
    fillColor = "red",
    stroke = FALSE,
    radius = 500,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```


### Asian population

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = asian_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_2(nh_asian_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_2,
    values = asian_map_labels$nh_asian_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  ) |> 
  addCircles(
    data = plants_map_labels,
    fillColor = "red",
    stroke = FALSE,
    radius = 500,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```

### Hispanic population

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = hispanic_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_3(hispanic_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_3,
    values = hispanic_map_labels$hispanic_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  ) |> 
  addCircles(
    data = plants_map_labels,
    fillColor = "red",
    stroke = FALSE,
    radius = 500,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```


### Other population

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = other_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_4(other_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_4,
    values = other_map_labels$other_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  ) |> 
  addCircles(
    data = plants_map_labels,
    fillColor = "red",
    stroke = FALSE,
    radius = 500,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```

### Income

```{r}
tx_income <- read_sf("interactive-data/income_counties.geojson")
```

```{r}
tx_income_labels <- tx_income |> 
  mutate(text_label = str_glue(
    "The median income in {Area} is {Median_HH_Income}"
  ))
```


```{r}
my_palette_5 <- colorNumeric(
  palette = "viridis",
  domain = tx_income_labels$State_HH_Inc_Percent
)
```

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = black_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_1(nh_black_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_1,
    values = black_map_labels$nh_black_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```
