---
title: "Data mapping"
---

## Goals of this notebook

I want to map the demographic data I found and cleaned in the first notebook.

I am following a tutorial from [here](https://rfortherestofus.com/courses/mapping/lessons/welcome-to-mapping-with-r).

## Setup

I am setting up the packages I need.

```{r}
library(leaflet)
library(tidyverse)
library(tidycensus)
library(janitor)
library(tigris)
library(scales)
library(sf)
library(tmap)
library(shiny)
```

## Getting the map started

Importing the data that I cleaned in the cleaning notebook.

```{r}
counties_map <- read_sf("interactive-data/race_counties.geojson")

counties_map
```


## Non-white category

Going to create a category that adds up Black, Asian and Hispanic populations and make that into a choropelth map.

```{r}
# nw_counties_map <- counties_map_wgs84 |> 
#   mutate(BIPOC_Pop = NH_Black + NH_Asian + NH_Other + Hispanic)
# 
# nw_counties_map
```

Now making the map.

```{r}
# my_palette <-
#   colorNumeric(
#     palette = "viridis",
#     domain = nw_counties_map$BIPOC_Pop
#   )
```

```{r}
# leaflet() |> 
#   addProviderTiles("CartoDB.Positron") |> 
#   addPolygons(
#     data = nw_counties_map,
#     fillColor = ~my_palette(BIPOC_Pop),
#     fillOpacity = 0.8,
#     color = "white",
#     weight = 0.75,
#     label = ~BIPOC_Pop
#   )
```

I am not a fan of how that looks, so I am uncommenting the code.

I am creating a labeled version to see how it would look.

First, creating the labels.

```{r}
# nw_counties_map_labels <- nw_counties_map |>
#   mutate(
#     text_label = str_glue(
#       "{BIPOC_Pop} of {Total} are considered as BIPOC in {Area}"
#     )
#   )
```

Making it into a map.

```{r}
# leaflet() |> 
#   addProviderTiles("CartoDB.Positron") |> 
#   addPolygons(
#     data = nw_counties_map_labels,
#     fillColor = ~my_palette(BIPOC_Pop),
#     fillOpacity = 0.8,
#     color = "white",
#     weight = 0.75,
#     popup = ~text_label
#   )
```

I don't think the labels make it better or more approachable. Uncommenting the code so it doesn't slow down the rest of the notebook.


## Four different maps

Making four different interactive maps for the different populations. Important note that the races are mutually exclusive groups and non-Hispanic other consists of individuals such as Alaska and American Natives, Pacific Islanders and all people listing two or more races, according to the Texas Demographic Center.

### Black population

Starting with Black population of Texas.

#### Percentages

Creating a column for the Black population percentage.

```{r}
black_map <- counties_map |> 
  mutate(nh_black_pct = NH_Black / Total) |>
  select(Area, nh_black_pct)

black_map
```

#### Static

Making a static map of the Black population in Texas.

```{r}
black_map |> 
  shift_geometry() |> 
  ggplot() +
  geom_sf(
    aes(
      fill = nh_black_pct,
      color = nh_black_pct
    )
  ) +
  theme_void() +
  theme(
    legend.position = "top",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.5, "cm"),
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16,
      margin = margin(
        b =0.25,
        unit = "cm"
      )
    )
  ) +
  scale_fill_viridis_c(limits = c(0,1),
                       labels = percent_format()) +
scale_color_viridis_c(limits = c(0,1),
                     labels = percent_format()) +
  labs(
    fill = NULL,
    color = NULL,
    title = "Percentage of Black Population Estimates for Texas Counties"
  )
```

#### Interactive

Making an interactive version.

Making the labels first.

```{r}
black_map_labels <- black_map |> 
  mutate(text_label = str_glue(
    "<b>{percent(nh_black_pct, accuracy = 0.1)}</b> of the population of <b>{Area}</b> identifies as Black"
  ))
```

Creating the color palette.

```{r}
my_palette_1 <- colorNumeric(
  palette = "viridis",
  domain = black_map_labels$nh_black_pct
)
```

Creating the map itself.

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = black_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_1(nh_black_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_1,
    values = black_map_labels$nh_black_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```


### Asian population

Now, doing the same for Asian population.

#### Percentage

Starting out with making a percentage of the Asian population in Texas.

```{r}
asian_map <- counties_map |> 
  mutate(nh_asian_pct = NH_Asian / Total) |>
  select(Area, nh_asian_pct)

asian_map
```

#### Static

Making the static map.

```{r}
asian_map |> 
  shift_geometry() |> 
  ggplot() +
  geom_sf(
    aes(
      fill = nh_asian_pct,
      color = nh_asian_pct
    )
  ) +
  theme_void() +
  theme(
    legend.position = "top",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.5, "cm"),
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16,
      margin = margin(
        b =0.25,
        unit = "cm"
      )
    )
  ) +
  scale_fill_viridis_c(limits = c(0,1),
                       labels = percent_format()) +
scale_color_viridis_c(limits = c(0,1),
                     labels = percent_format()) +
  labs(
    fill = NULL,
    color = NULL,
    title = "Percentage of Asian Population Estimates for Texas Counties"
  )
```

#### Interactive

Making an interactive version.

Starting with making the labels.

```{r}
asian_map_labels <- asian_map |> 
  mutate(text_label = str_glue(
    "<b>{percent(nh_asian_pct, accuracy = 0.1)}</b> of the population of <b>{Area}</b> identifies as Asian"
  ))
```


Making a color palette.

```{r}
my_palette_2 <- colorNumeric(
  palette = "viridis",
  domain = asian_map_labels$nh_asian_pct
)
```


Creating the map itself.

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = asian_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_2(nh_asian_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_2,
    values = asian_map_labels$nh_asian_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```

### Hispanic

Applying same Process for Hispanic population.

#### Percentage

Starting with making the percentage of the Hispanic population in Texas.

```{r}
hispanic_map <- counties_map |> 
  mutate(hispanic_pct = Hispanic / Total) |>
  select(Area, hispanic_pct)

hispanic_map
```

#### Static

Making the static map.

```{r}
hispanic_map |> 
  shift_geometry() |> 
  ggplot() +
  geom_sf(
    aes(
      fill = hispanic_pct,
      color = hispanic_pct
    )
  ) +
  theme_void() +
  theme(
    legend.position = "top",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.5, "cm"),
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16,
      margin = margin(
        b =0.25,
        unit = "cm"
      )
    )
  ) +
  scale_fill_viridis_c(limits = c(0,1),
                       labels = percent_format()) +
scale_color_viridis_c(limits = c(0,1),
                     labels = percent_format()) +
  labs(
    fill = NULL,
    color = NULL,
    title = "Percentage of Hispanic Population Estimates for Texas Counties"
  )
```

#### Interactive

Making an interactive version.

Making the labels first.

```{r}
hispanic_map_labels <- hispanic_map |> 
  mutate(text_label = str_glue(
    "<b>{percent(hispanic_pct, accuracy = 0.1)}</b> of the population of <b>{Area}</b> identifies as Hispanic"
  ))
```

Now creating a color palette.

```{r}
my_palette_3 <- colorNumeric(
  palette = "viridis",
  domain = hispanic_map_labels$hispanic_pct
)
```

Making the map itself.

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = hispanic_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_3(hispanic_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_3,
    values = hispanic_map_labels$hispanic_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```

### Other

Finishing it off with other population, which, as mentioned above, includes Native Americans and people with multiple races.

#### Percentages

Making the percentage of population that identifies as other in Texas.

```{r}
other_map <- counties_map |> 
  mutate(other_pct = NH_Other / Total) |>
  select(Area, other_pct)

other_map
```

#### Static

Making the static map.

```{r}
other_map |> 
  shift_geometry() |> 
  ggplot() +
  geom_sf(
    aes(
      fill = other_pct,
      color = other_pct
    )
  ) +
  theme_void() +
  theme(
    legend.position = "top",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.5, "cm"),
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16,
      margin = margin(
        b =0.25,
        unit = "cm"
      )
    )
  ) +
  scale_fill_viridis_c(limits = c(0,1),
                       labels = percent_format()) +
scale_color_viridis_c(limits = c(0,1),
                     labels = percent_format()) +
  labs(
    fill = NULL,
    color = NULL,
    title = "Percentage of Individuals Identifying as Other Estimates for Texas Counties"
  )
```

#### Interactive

Making an interactive version.

Making the labels first.

```{r}
other_map_labels <- other_map |> 
  mutate(text_label = str_glue(
    "<b>{percent(other_pct, accuracy = 0.1)}</b> of the population of <b>{Area}</b> identifies as Other"
  ))
```

Creating a color palette.

```{r}
my_palette_4 <- colorNumeric(
  palette = "viridis",
  domain = other_map_labels$other_pct
)
```

Now making the map itself.

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = other_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_4(other_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_4,
    values = other_map_labels$other_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  )
```

## Plant data

Mapping the plant data as locations.

Starting out with a static map.

```{r}
plants_map <- read_sf("interactive-data/tx-plants-data.geojson")

plants_map
```

Using a package to import the geometry of the state of Texas to create the base of the static map.

```{r}
texas <- map_data('county', region = 'texas')

texas
```

Importing the rds version of the Texas plants.

```{r}
tx_plants <- read_rds("interactive-data/tx-plants-data.rds")
```

Creating the static map.

Starting out with the base layer.

```{r}
texas_map <- ggplot(data = texas) +
  geom_polygon(aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  coord_fixed(1.3)

texas_map
```

Putting the plant locations on the map.

```{r}
texas_map +
  geom_point(data = tx_plants, aes(x = Longitude, y = Latitude), color = "blue", size = 0.75) +
  theme_void()
```

Done with static, making an interactive version.

Creating labels for the map.

```{r}
plants_map_labels <- plants_map |> 
  mutate(text_label = str_glue(
    "{Project} developed by {Company}"
  ))
```

Creating the interactive map.

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addCircles(
    data = plants_map_labels,
    fillColor = "blue",
    stroke = FALSE,
    radius = 500,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```

## Adding both sets of data on a map

### Black population

```{r}
leaflet() |> 
   addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = black_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_1(nh_black_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_1,
    values = black_map_labels$nh_black_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  ) |> 
  addCircles(
    data = plants_map_labels,
    fillColor = "red",
    stroke = FALSE,
    radius = 500,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```


### Asian population

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = asian_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_2(nh_asian_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_2,
    values = asian_map_labels$nh_asian_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  ) |> 
  addCircles(
    data = plants_map_labels,
    fillColor = "red",
    stroke = FALSE,
    radius = 500,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```

### Hispanic population

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = hispanic_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_3(hispanic_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_3,
    values = hispanic_map_labels$hispanic_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  ) |> 
  addCircles(
    data = plants_map_labels,
    fillColor = "red",
    stroke = FALSE,
    radius = 500,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```


### Other population

```{r}
leaflet() |> 
  addProviderTiles("CartoDB.Positron") |> 
  addPolygons(
    data = other_map_labels,
    weight = 0.75,
    fillColor = ~my_palette_4(other_pct),
    fillOpacity = 0.8,
    color = "white",
    popup = ~text_label
  ) |> 
  addLegend(
    position = "topright",
    pal = my_palette_4,
    values = other_map_labels$other_pct,
    labFormat = labelFormat(
      transform = function(x) sort(x * 100),
      suffix = "%"
    ),
    opacity = 0.8
  ) |> 
  addCircles(
    data = plants_map_labels,
    fillColor = "red",
    stroke = FALSE,
    radius = 500,
    fillOpacity = 1,
    popup = ~text_label
  ) 
```

### Income

Importing the median household income per county.

```{r}
tx_income <- read_sf("interactive-data/income_counties.geojson")
```

I know that some of the data is character instead of a number, so I am fixing that here.

```{r}
tx_income |> 
  mutate(across(c(Median_HH_Income, State_HH_Inc_Percent), parse_number),
         HH_Inc_Perc = State_HH_Inc_Percent / 100)
```

Now creating the labels for the map.

```{r}
tx_income_labels <- tx_income |> 
  mutate(text_label = str_glue(
    "The median income in {Area} is {Median_HH_Income}"
  ))
```

Creating the color palette now.

```{r}
my_palette_5 <- colorNumeric(
  palette = "viridis",
  domain = tx_income_labels$HH_Inc_Perc
)
```

